---
title: "MasterControl Lead Progression Modeling"
author: "Adam Edwards, An Lee, Brian Frerichs, Kun Joo Cho"
date: "2026-03-13"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    code_folding: hide
    number_sections: true
  output:
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

# Introduction

## Business Problem

MasterControl provides quality and manufacturing management software to life sciences companies. Their two main products are:

- **Qx (Quality solutions)**: 20+ years in market, ~19.7% lead progression rate
- **Mx (Manufacturing solutions)**: ~4 years in market, ~12.7% lead progression rate

Sales resources are limited. The goal is to build a predictive model that identifies which leads are most likely to progress, so that sales teams can prioritize the highest-potential opportunities — especially for the underperforming Mx product.

## Analytic Problem

We aim to build a binary classification model predicting whether a lead progresses to a successful sales stage (SQL, SQO, or Won). 

- **Target variable**: `next_stage_c` (binarized: 1 = progressed, 0 = not progressed)
- **Baseline rate**: ~18.45% overall progression
- **Key challenge**: Mx product significantly underperforms Qx; we want to understand and model this gap

# Data Preparation

## Load Libraries

```{r libraries}
library(tidyverse)
library(tidymodels)
library(themis)
library(vip)
library(ranger)
library(xgboost)
library(glmnet)
```

## Load & Rename

```{r Load-Data}
df_raw <- read_csv("Mastercontrol QAL Performance.csv")

df <- df_raw %>%
  rename(
    site_function = 'acct primary site function',
    mfg_model     = 'acct manufacturing model',
    industry      = 'acct target industry',
    lead_title    = 'contact/lead title',
    lead_id       = 'contact/lead id',
    next_stage    = 'next_stage__c',
    territory     = 'acct territory rollup',
    tier          = 'acct tier rollup',
    channel       = 'last tactic campaign channel',
    cohort_date   = 'qal cohort date'
  )

```

## Create Binary Target

```{r Target}
df_model <- df %>%
  filter(!is.na(next_stage)) %>%
  mutate(
    progressed = factor(
      if_else(next_stage %in% c("SQL", "SQO", "Won"), "Yes", "No"),
      levels = c("Yes", "No")
    )
  )

table(df_model$progressed)
prop.table(table(df_model$progressed))
```

## Feature Engineering

```{r Features Engineering}
df_model <- df_model %>%
  mutate(
    cohort_date  = as.Date(cohort_date, format = "%m/%d/%Y"),
    cohort_month = month(cohort_date),
    cohort_qtr   = quarter(cohort_date),
    cohort_year  = year(cohort_date),

    title_upper = str_to_upper(lead_title),

    seniority = case_when(
      str_detect(title_upper, "CEO|CTO|COO|CFO|CHIEF|PRESIDENT|OWNER|FOUNDER") ~ "C-Level",
      str_detect(title_upper, "\\bVP\\b|VICE PRESIDENT")                       ~ "VP",
      str_detect(title_upper, "DIRECTOR|DIR\\b")                               ~ "Director",
      str_detect(title_upper, "MANAGER|MGR\\b")                                ~ "Manager",
      str_detect(title_upper, "SENIOR|SR\\.|LEAD\\b|PRINCIPAL")                ~ "Senior/Lead",
      str_detect(title_upper, "SPECIALIST|ANALYST|COORDINATOR|ENGINEER|SCIENTIST") ~ "Technical",
      !is.na(lead_title) ~ "Other",
      TRUE ~ "Unknown"
    ),

    func_area = case_when(
      str_detect(title_upper, "QUALITY|QA\\b|QC\\b|COMPLIANCE|VALIDATION") ~ "Quality",
      str_detect(title_upper, "REGULAT|RA\\b|AFFAIRS")                      ~ "Regulatory",
      str_detect(title_upper, "MANUFACTUR|OPERAT|PRODUC|SUPPLY CHAIN")      ~ "Operations/Mfg",
      str_detect(title_upper, "IT\\b|INFORMATION|SYSTEM|TECHNOLOG|DATA")    ~ "IT/Systems",
      str_detect(title_upper, "RESEARCH|R&D|DEVELOP|SCIENTIST|CLINICAL")    ~ "R&D",
      str_detect(title_upper, "ENGINEER")                                    ~ "Engineering",
      !is.na(lead_title) ~ "Other",
      TRUE ~ "Unknown"
    ),

    kw_quality_cluster = as.integer(str_detect(title_upper, "QUALITY|\\bQA\\b|\\bQC\\b|COMPLIANCE|VALIDATION")),
    kw_mfg_cluster     = as.integer(str_detect(title_upper, "MANUFACTUR|OPERAT|PRODUC|SUPPLY CHAIN")),

    is_senior_decision_maker = as.integer(seniority %in% c("C-Level", "VP")),

    is_quality_or_ops = as.integer(func_area %in% c("Quality", "Operations/Mfg")),

    high_intent    = as.integer(str_detect(Priority, "P1 -|Priority 1")),
    is_pricing_req = as.integer(str_detect(Priority, "Pricing")),
    is_demo_req    = as.integer(str_detect(Priority, "Demo|Video Demo|Live Demo")),

    organic_channel = as.integer(channel %in% c("SEO", "Direct/Inbound")),

    is_americas = as.integer(territory == "Americas"),

    is_small_med = as.integer(tier %in% c("Small", "Medium")),

    is_target_industry_mx = as.integer(industry %in% c("Medical Device", "Blood & Biologics")),

    site_function = replace_na(site_function, "Missing"),
    mfg_model     = replace_na(mfg_model, "Missing"),
    industry      = replace_na(industry, "Unknown"),
    territory     = replace_na(territory, "Unknown"),
    tier          = replace_na(tier, "Unknown"),
    channel       = replace_na(channel, "Unknown"),

    mfg_model_grp  = fct_lump_n(factor(mfg_model), n = 10, other_level = "Other"),
    channel_grp    = fct_lump_n(factor(channel),    n = 8,  other_level = "Other"),
    priority_grp   = fct_lump_n(factor(Priority),   n = 8,  other_level = "Other"),
    industry_grp   = factor(industry),
    territory_grp  = factor(territory),
    tier_grp       = factor(tier),
    seniority_grp  = factor(seniority),
    func_area_grp  = factor(func_area),
    product_grp    = factor(solution_rollup)
  )
```

```{r Correlation & Progression}
cor_check <- df_model %>%
  mutate(across(c(kw_quality_cluster, kw_mfg_cluster,
                  high_intent, organic_channel), as.integer)) %>%
  select(kw_quality_cluster, kw_mfg_cluster,
         high_intent, organic_channel) %>%
  cor(use = "complete.obs")

print(cor_check)

df_model %>%
  group_by(func_area_grp, product_grp) %>%
  summarise(n = n(),
            prog = mean(progressed == "Yes"),
            .groups = "drop") %>%
  filter(product_grp == "Mx", n >= 30) %>%
  arrange(desc(prog))


month_prog <- df_model %>%
  group_by(cohort_month, product_grp) %>%
  summarise(n = n(),
            prog_rate = mean(progressed == "Yes"),
            .groups = "drop") %>%
  filter(product_grp %in% c("Mx", "Qx"), n >= 20)

ggplot(month_prog,
       aes(x = cohort_month, y = prog_rate,
           color = product_grp, group = product_grp)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = 1:12,
    labels = c("Jan","Feb","Mar","Apr","May","Jun",
               "Jul","Aug","Sep","Oct","Nov","Dec")) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Monthly Progression Rate (Mx vs Qx)",
      x = "Lead Cohort Month", y = "Progression Rate", color = "Product") +
  theme_minimal()

df_model %>%
  filter(product_grp == "Mx") %>%
  group_by(cohort_year) %>%
  summarise(n = n(),
            prog_rate = mean(progressed == "Yes"),
            .groups = "drop")
```

```{r Site Function Tiering}
site_rates <- df_model %>%
  group_by(site_function) %>%
  summarise(n = n(), prog_rate = mean(progressed == "Yes"), .groups = "drop") %>%
  filter(n >= 20) %>%
  mutate(site_func_tier = case_when(
    prog_rate >= 0.30 ~ "High",
    prog_rate >= 0.18 ~ "Mid",
    TRUE              ~ "Low"
  )) %>%
  select(site_function, site_func_tier)

df_model <- df_model %>%
  left_join(site_rates, by = "site_function") %>%
  mutate(
    site_func_tier = replace_na(site_func_tier, "Unknown"),
    site_func_tier = factor(site_func_tier, levels = c("High", "Mid", "Low", "Unknown"))
  )
```

## Select Final Features

```{r Select-Features}
df_final <- df_model %>%
  select(
    progressed,
    product_grp,
    site_func_tier,
    industry_grp, is_target_industry_mx,
    mfg_model_grp,
    tier_grp, is_small_med,
    territory_grp, is_americas,
    seniority_grp, is_senior_decision_maker,
    func_area_grp, is_quality_or_ops,
    kw_quality_cluster, kw_mfg_cluster,
    priority_grp, high_intent, is_pricing_req, is_demo_req,
    channel_grp, organic_channel,
    cohort_month, cohort_qtr, cohort_year
  )
```

```{r Eefine Clustering}
df_model <- df_model %>%
  mutate(
    kw_qa           = as.integer(str_detect(title_upper, "\\bQA\\b|\\bQC\\b")),
    kw_compliance   = as.integer(str_detect(title_upper, "COMPLIANCE|COMPLI")),
    kw_validation   = as.integer(str_detect(title_upper, "VALIDATION|VALIDAT")),
    kw_operations   = as.integer(str_detect(title_upper, "OPERAT")),
    kw_supply_chain = as.integer(str_detect(title_upper, "SUPPLY CHAIN")),
    kw_quality_cluster = as.integer(kw_qa | kw_compliance | kw_validation),
    kw_mfg_cluster     = as.integer(kw_operations | kw_supply_chain)
  )

kw_cols <- c("kw_qa", "kw_compliance", "kw_validation",
             "kw_operations", "kw_quality_cluster", "kw_mfg_cluster")

map_dfr(kw_cols, function(col) {
  df_model %>%
    group_by(flag = .data[[col]]) %>%
    summarise(n = n(), prog_rate = mean(progressed == "Yes"), .groups = "drop") %>%
    mutate(keyword = col)
}) %>%
  filter(flag == 1) %>%
  select(keyword, n, prog_rate) %>%
  arrange(desc(prog_rate))
```

```{r Clean Site Tiers}

site_rates <- df_model %>%
  group_by(site_function) %>%
  summarise(n = n(), prog_rate = mean(progressed == "Yes"), .groups = "drop") %>%
  filter(n >= 20)

site_groups <- site_rates %>%
  mutate(
    site_func_tier = case_when(
      prog_rate >= 0.30 ~ "High",
      prog_rate >= 0.18 ~ "Mid",
      TRUE              ~ "Low"
    )
  ) %>%
  select(site_function, site_func_tier)

# Remove site_func_tier columns only if they exist
df_model <- df_model %>%
  select(-any_of(c("site_func_tier", "site_func_tier.x", "site_func_tier.y")))

# Compute progression rate per site function (min 20 samples)
site_rates <- df_model %>%
  group_by(site_function) %>%
  summarise(n = n(), prog_rate = mean(progressed == "Yes"), .groups = "drop") %>%
  filter(n >= 20)

site_groups <- site_rates %>%
  mutate(
    site_func_tier = case_when(
      prog_rate >= 0.30 ~ "High",
      prog_rate >= 0.18 ~ "Mid",
      TRUE              ~ "Low"
    )
  ) %>%
  select(site_function, site_func_tier)

df_model <- df_model %>%
  left_join(site_groups, by = "site_function") %>%
  mutate(
    site_func_tier = replace_na(site_func_tier, "Unknown"),
    site_func_tier = factor(site_func_tier, levels = c("High", "Mid", "Low", "Unknown"))
  )

# Verify
df_model %>%
  group_by(site_func_tier) %>%
  summarise(n = n(), prog_rate = mean(progressed == "Yes"), .groups = "drop")
```

# Plots of MX Relationships

```{r}
# Keep only Mx leads from the processed dataset
mx <- df_final %>% filter(product_grp == "Mx")

# Utility summarizer (handles factor target)
rate_by <- function(data, group_var, min_n = 30) {
  data %>%
    group_by({{ group_var }}) %>%
    summarize(
      n = n(),
      rate = mean(progressed == "Yes"),  # works with factor
      .groups = "drop"
    ) %>%
    filter(n >= min_n)
}

# --- Persona (function) ---
rate_by(mx, func_area_grp) %>%
  ggplot(aes(fct_reorder(func_area_grp, rate), rate)) +
  geom_col(fill = "#3C8DBC") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Mx Progression by Title Function",
       x = "Function", y = "Progression Rate") +
  theme_minimal()
```

```{r Persona (seniority)}

rate_by(mx, seniority_grp) %>%
  ggplot(aes(fct_reorder(seniority_grp, rate), rate)) +
  geom_col(fill = "#8172B3") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Mx Progression by Seniority") +
  theme_minimal()
```

```{r Industry}

# Ensure mx uses the processed dataset
mx <- df_final %>% filter(product_grp == "Mx")

rate_by(mx, industry_grp) %>%
  ggplot(aes(fct_reorder(industry_grp, rate), rate)) +
  geom_col(fill = "#4C72B0") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Mx Progression by Industry") +
  theme_minimal()
```

```{r Channel}

# Ensure mx is from the correct dataset (df_final or df_model)
mx <- df_final %>% filter(product_grp == "Mx")

# Define rate_by if not already defined
rate_by <- function(data, group_var, min_n = 30) {
  data %>%
    group_by({{ group_var }}) %>%
    summarize(
      n = n(),
      rate = mean(progressed == "Yes"),
      .groups = "drop"
    ) %>%
    filter(n >= min_n)
}

# Plot by channel_grp
rate_by(mx, channel_grp) %>%
  ggplot(aes(fct_reorder(channel_grp, rate), rate)) +
  geom_col(fill = "#70AD47") +
  coord_flip() +
  labs(title = "Mx Progression by Channel",
       y = "Progression Rate") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

```{r Priority}
mx_raw <- df_model %>% filter(product_grp == "Mx")
rate_by(mx_raw, Priority) %>%
  ggplot(aes(fct_reorder(Priority, rate), rate)) +
  geom_col(fill = "#ED7D31") +
  coord_flip() +
  labs(title = "Mx Progression by Lead Priority",
       y = "Progression Rate") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

# Plots of MX & Qx Relationships

```{r}
qx_mx <- df_model %>% filter(product_grp %in% c("Qx", "Mx"))

# Utility summarizer (handles factor target)
rate_by <- function(data, group_var, min_n = 30) {
  data %>%
    group_by({{ group_var }}) %>%
    summarize(n = n(), rate = mean(progressed == "Yes"), .groups = "drop") %>%
    filter(n >= min_n)
}

# --- Persona (function) ---
rate_by(qx_mx, func_area_grp) %>%
  ggplot(aes(fct_reorder(func_area_grp, rate), rate)) +
  geom_col(fill = "#3C8DBC") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Qx & Mx Progression by Title Function",
       x = "Function", y = "Progression Rate") +
  theme_minimal()
```

```{r 2nd Persona (seniority)}

rate_by(qx_mx, seniority) %>%
  ggplot(aes(fct_reorder(seniority, rate), rate)) +
  geom_col(fill = "#8172B3") +
  coord_flip() +
  scale_y_continuous(labels = percent) +
  labs(title = "qx_mx Progression by Seniority") +
  theme_minimal()
```

```{r 2nd Industry}

rate_by(qx_mx, industry_grp) %>%
  ggplot(aes(fct_reorder(industry_grp, rate), rate)) +
  geom_col(fill = "#4C72B0") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Qx & Mx Progression by Industry",
       x = "Industry",
       y = "Progression Rate") +
  theme_minimal()
```

```{r Manufacturing Model}
rate_by(qx_mx, mfg_model_grp) %>%
  ggplot(aes(fct_reorder(mfg_model_grp, rate), rate)) +
  geom_col(fill = "#5B9BD5") +
  coord_flip() +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Qx & Mx Progression by Manufacturing Model",
       x = "Manufacturing Model",
       y = "Progression Rate") +
  theme_minimal()
```

```{r 2nd Channel}
rate_by(qx_mx, channel_grp) %>%
  ggplot(aes(fct_reorder(channel_grp, rate), rate)) +
  geom_col(fill = "#70AD47") +
  coord_flip() +
  labs(title = "Qx & Mx Progression by Channel",
       x = "Channel",
       y = "Progression Rate") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

```{r 2nd Priority}

rate_by(qx_mx, Priority) %>%  # Note: capital P
  ggplot(aes(fct_reorder(Priority, rate), rate)) +
  geom_col(fill = "#ED7D31") +
  coord_flip() +
  labs(title = "Qx & Mx Progression by Lead Priority",
       x = "Priority",
       y = "Progression Rate") +
  scale_y_continuous(labels = scales::percent) +
  theme_minimal()
```

# Modeling

## Train/Test Split & CV

```{r Split}
set.seed(42)
splits   <- initial_split(df_final, prop = 0.8, strata = progressed)
train    <- training(splits)
test     <- testing(splits)
cv_folds <- vfold_cv(train, v = 5, strata = progressed)
```

## Preprocessing Recipe

```{r Recipe}
rec <- recipe(progressed ~ ., data = train) %>%
  step_impute_median(all_numeric_predictors()) %>%
  step_impute_mode(all_nominal_predictors()) %>%
  step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
  step_zv(all_predictors()) %>%
  step_normalize(all_numeric_predictors()) %>%
  step_downsample(progressed, under_ratio = 1, seed = 42)
```

## Benchmark

```{r Benchmark}
# Majority-class classifier: predict "No"
cat("Benchmark accuracy (always No):", mean(test$progressed == "No"), "\n")
cat("Benchmark AUC: 0.50 (random classifier)\n")
```

```{r Model Benchmark}
# Helper function: train and evaluate a Random Forest for a single product
run_product_model <- function(data, product_name) {

  df_sub <- data %>% filter(product_grp == product_name)

  set.seed(42)
  splits <- initial_split(df_sub, prop = 0.8, strata = progressed)
  train  <- training(splits)
  test   <- testing(splits)
  folds  <- vfold_cv(train, v = 5, strata = progressed)

  rec <- recipe(progressed ~ ., data = train) %>%
    step_impute_median(all_numeric_predictors()) %>%
    step_impute_mode(all_nominal_predictors()) %>%
    step_dummy(all_nominal_predictors(), one_hot = FALSE) %>%
    step_zv(all_predictors()) %>%
    step_normalize(all_numeric_predictors()) %>%
    step_downsample(progressed, under_ratio = 1, seed = 42)

  rf_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 500) %>%
    set_engine("ranger", importance = "impurity") %>%
    set_mode("classification")

  set.seed(42)
  rf_tune <- tune_grid(
    workflow() %>% add_recipe(rec) %>% add_model(rf_spec),
    resamples = folds,
    grid = grid_regular(mtry(range = c(3, 8)), min_n(range = c(5, 30)), levels = 4),
    metrics = metric_set(roc_auc)
  )

  best_rf  <- select_best(rf_tune, metric = "roc_auc")
  final_wf <- finalize_workflow(workflow() %>% add_recipe(rec) %>% add_model(rf_spec), best_rf)
  final_fit <- last_fit(final_wf, splits, metrics = metric_set(roc_auc, f_meas))

  list(
    product   = product_name,
    final_wf  = final_wf,
    final_fit = final_fit,
    train     = train,
    cv_auc    = show_best(rf_tune, metric = "roc_auc", n = 1)$mean,
    test_auc  = collect_metrics(final_fit) %>% filter(.metric == "roc_auc") %>% pull(.estimate)
  )
}

mx_result <- run_product_model(df_final, "Mx")
qx_result <- run_product_model(df_final, "Qx")

# Product-level performance summary
tibble(
  Product  = c("Mx", "Qx"),
  CV_AUC   = c(mx_result$cv_auc,  qx_result$cv_auc),
  Test_AUC = c(mx_result$test_auc, qx_result$test_auc)
)
```

## Logistic Regression

```{r logistic}
lr_spec <- logistic_reg(penalty = tune(), mixture = tune()) %>%
  set_engine("glmnet") %>%
  set_mode("classification")

lr_wf <- workflow() %>% add_recipe(rec) %>% add_model(lr_spec)

set.seed(42)
lr_tune <- tune_grid(
  lr_wf,
  resamples = cv_folds,
  grid      = grid_regular(penalty(), mixture(), levels = 5),
  metrics   = metric_set(roc_auc)
)

best_lr <- select_best(lr_tune, metric = "roc_auc")
cat("Best LR - CV AUC:", show_best(lr_tune, metric = "roc_auc", n = 1)$mean, "\n")
```

## Random Forest

```{r random-forest}
rf_spec <- rand_forest(mtry = tune(), min_n = tune(), trees = 500) %>%
  set_engine("ranger", importance = "impurity", max.depth = 10) %>%
  set_mode("classification")

rf_wf <- workflow() %>% add_recipe(rec) %>% add_model(rf_spec)

set.seed(42)
rf_tune <- tune_grid(
  rf_wf,
  resamples = cv_folds,
  grid      = grid_regular(mtry(range = c(3, 10)), min_n(range = c(10, 50)), levels = 4),
  metrics   = metric_set(roc_auc)
)

best_rf <- select_best(rf_tune, metric = "roc_auc")
cat("Best RF - CV AUC:", show_best(rf_tune, metric = "roc_auc", n = 1)$mean, "\n")
```

## XGBoost

```{r xgboost}
xgb_spec <- boost_tree(
  trees = tune(), tree_depth = tune(),
  learn_rate = tune(), min_n = tune(),
  sample_size = tune()
) %>%
  set_engine("xgboost") %>%
  set_mode("classification")

xgb_wf <- workflow() %>% add_recipe(rec) %>% add_model(xgb_spec)

set.seed(42)
xgb_tune <- tune_grid(
  xgb_wf,
  resamples = cv_folds,
  grid = grid_latin_hypercube(
    trees(range = c(100, 600)), tree_depth(range = c(3, 8)),
    learn_rate(range = c(-3, -1)), min_n(range = c(10, 50)),
    sample_size = sample_prop(),
    size = 25
  ),
  metrics = metric_set(roc_auc)
)

best_xgb <- select_best(xgb_tune, metric = "roc_auc")
cat("Best XGB - CV AUC:", show_best(xgb_tune, metric = "roc_auc", n = 1)$mean, "\n")
```


# Model Performance

## CV AUC Comparison

```{r Compare}
tibble(
  Model  = c("Logistic Regression", "Random Forest", "XGBoost"),
  CV_AUC = c(
    show_best(lr_tune,  metric = "roc_auc", n = 1)$mean,
    show_best(rf_tune,  metric = "roc_auc", n = 1)$mean,
    show_best(xgb_tune, metric = "roc_auc", n = 1)$mean
  )
) %>% arrange(desc(CV_AUC))
```

## Final Model: Test Set Evaluation

```{r Final-Eval}
# Random Forest is highest AUC
final_wf  <- finalize_workflow(rf_wf, best_rf)
final_fit <- last_fit(final_wf, splits, metrics = metric_set(roc_auc, accuracy, f_meas))

collect_metrics(final_fit)
```

## In-Sample vs Out-of-Sample AUC

```{r In-vs-Out}
# Diagnose overfitting: compare train vs test AUC and flag if gap > 0.05
diagnose_overfit <- function(result) {
  final_model <- fit(result$final_wf, data = result$train)
  train_preds <- augment(final_model, new_data = result$train)
  test_preds  <- collect_predictions(result$final_fit)

  in_auc  <- roc_auc(train_preds, truth = progressed,
                     .pred_Yes, event_level = "first")$.estimate
  out_auc <- result$test_auc

  cat(sprintf(
    "[%s] Train AUC: %.4f | Test AUC: %.4f | Gap: %.4f %s\n",
    result$product, in_auc, out_auc, in_auc - out_auc,
    ifelse(in_auc - out_auc > 0.05, "<- WARNING: overfitting", "OK")
  ))
}

diagnose_overfit(mx_result)
diagnose_overfit(qx_result)

# Find optimal threshold under three business scenarios:
#   (A) Maximize F1
#   (B) Precision >= 0.50 (limited sales capacity, avoid wasted effort)
#   (C) Recall    >= 0.70 (do not miss high-potential leads)
optimize_threshold <- function(result) {
  preds <- collect_predictions(result$final_fit)

  map_dfr(seq(0.25, 0.75, by = 0.05), function(t) {
    p <- preds %>%
      mutate(.pred_class = factor(if_else(.pred_Yes >= t, "Yes", "No"),
                                  levels = c("Yes", "No")))
    tibble(
      threshold = t,
      f1        = f_meas(p,    truth = progressed, estimate = .pred_class)$.estimate,
      precision = precision(p, truth = progressed, estimate = .pred_class)$.estimate,
      recall    = recall(p,    truth = progressed, estimate = .pred_class)$.estimate
    )
  }) -> tbl

  cat(sprintf("\n--- %s threshold optimization ---\n", result$product))
  cat(sprintf("(A) Max F1:          t=%.2f | F1=%.3f | Prec=%.3f | Rec=%.3f\n",
    { r <- slice_max(tbl, f1, n=1); r$threshold }, { r <- slice_max(tbl, f1, n=1); r$f1 },
    { r <- slice_max(tbl, f1, n=1); r$precision }, { r <- slice_max(tbl, f1, n=1); r$recall }))
  cat(sprintf("(B) Prec >= 0.50:    t=%.2f | F1=%.3f | Prec=%.3f | Rec=%.3f\n",
    { r <- filter(tbl, precision >= 0.50) %>% slice_max(f1, n=1); r$threshold },
    { r <- filter(tbl, precision >= 0.50) %>% slice_max(f1, n=1); r$f1 },
    { r <- filter(tbl, precision >= 0.50) %>% slice_max(f1, n=1); r$precision },
    { r <- filter(tbl, precision >= 0.50) %>% slice_max(f1, n=1); r$recall }))
  cat(sprintf("(C) Recall >= 0.70:  t=%.2f | F1=%.3f | Prec=%.3f | Rec=%.3f\n",
    { r <- filter(tbl, recall >= 0.70) %>% slice_max(f1, n=1); r$threshold },
    { r <- filter(tbl, recall >= 0.70) %>% slice_max(f1, n=1); r$f1 },
    { r <- filter(tbl, recall >= 0.70) %>% slice_max(f1, n=1); r$precision },
    { r <- filter(tbl, recall >= 0.70) %>% slice_max(f1, n=1); r$recall }))
}

optimize_threshold(mx_result)
optimize_threshold(qx_result)
```

```{r Fit}
# Rebuild test_preds from the combined model final_fit
test_preds <- collect_predictions(final_fit)
```

```{r Threshold}
thresholds <- seq(0.3, 0.7, by = 0.05)

threshold_results <- map_dfr(thresholds, function(t) {
  preds <- test_preds %>%
    mutate(.pred_class = factor(
      if_else(.pred_Yes >= t, "Yes", "No"),
      levels = c("Yes", "No")
    ))
  tibble(
    threshold = t,
    f1        = f_meas(preds,    truth = progressed, estimate = .pred_class)$.estimate,
    precision = precision(preds, truth = progressed, estimate = .pred_class)$.estimate,
    recall    = recall(preds,    truth = progressed, estimate = .pred_class)$.estimate
  )
})

threshold_results
```

```{r Threshold-apply}
best_threshold <- threshold_results %>%
  arrange(desc(f1)) %>%
  dplyr::slice(1) %>%
  pull(threshold)

test_preds_adj <- test_preds %>%
  mutate(.pred_class = factor(
    if_else(.pred_Yes >= best_threshold, "Yes", "No"),
    levels = c("Yes", "No")
  ))
```


## Confusion Matrix

```{r Confusion}
conf_mat(test_preds_adj, truth = progressed, estimate = .pred_class)
```

```{r AUC & ROC}
# Restore out_auc for the combined model ROC plot
out_auc <- collect_metrics(final_fit) %>% 
  filter(.metric == "roc_auc") %>% 
  pull(.estimate)
```

## ROC Curve

```{r ROC}
roc_curve(test_preds, truth = progressed, .pred_Yes, event_level = "first") %>%
  ggplot(aes(x = 1 - specificity, y = sensitivity)) +
  geom_path() +
  geom_abline(lty = 2) +
  labs(title = paste0("ROC Curve (AUC = ", round(out_auc, 3), ")")) +
  theme_minimal()
```

```{r Final Model}
final_model <- fit(final_wf, data = train)
```

## Variable Importance

```{r Vip}
final_model %>%
  extract_fit_parsnip() %>%
  vip(num_features = 15)

# Extract top-N variable importance per product and compare side by side
extract_vip <- function(result, top_n = 15) {
  fit(result$final_wf, data = result$train) %>%
    extract_fit_parsnip() %>%
    vi() %>%
    slice_max(Importance, n = top_n) %>%
    mutate(Product = result$product)
}

bind_rows(extract_vip(mx_result), extract_vip(qx_result)) %>%
  select(Product, Variable, Importance) %>%
  pivot_wider(names_from = Product, values_from = Importance) %>%
  arrange(desc(coalesce(Mx, 0))) %>%
  print(n = 20)
```


# Results

We tested three models. Random Forest was the best with a test AUC of 0.855, followed by XGBoost (0.838) and Logistic Regression (0.829). All three scored well above the 0.50 baseline, so the data has clear, usable patterns.
We also built separate models for Mx and Qx. Both scored above 0.83, meaning we can reliably score leads for both products.
At the chosen threshold of 0.60, the model catches 72% of leads that would progress. Teams with less capacity can raise the threshold to get a shorter, cleaner list — but they will miss more real leads.

### Business Insights

The single biggest signal is what a lead does, not who they are. Pricing page visits convert at 58% for Mx and 50% for Qx — far above the 18.5% average. Demo requests follow at 25–32%. These leads should be called first, every time.
Mx and Qx need different targeting. Mx works best with small-to-mid sized Medical Device and Biologics companies in the Americas. Quality and Engineering roles convert the highest at around 16%. Focusing here is the fastest way to close the gap between Mx (12.9%) and Qx (20.3%).

On channels, SEO and Direct/Inbound convert at 24–31% for both products. Online Ads for Mx sit at only 19%. Organic traffic brings better leads, so shifting budget there makes sense.
Lastly, 40% of records are missing key fields like job title and site function. Leads with missing data convert much less. Making these fields required at sign-up is a simple fix with a real impact.
